Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-12-06 Ben_Manuscripts/structure_paper/Outline.tex
Turning outline into draft - intro text nearly done

  Despite having structural data, there is still information which
  experiment cannot definitively answer. There are several key
  questions that we intend to answer 
%MRS3: we can't ``answer'' for sure, since it's a model, and may omit things. Could put ``several key questions we will investigate'' or ``hope to provide evidence for or against'', or something like that.
which will be laid out and numbered
  in subsequent paragraphs.

@@

  Monomers in the Col\textsubscript{h} system are theorized to be partitioned
  into stacked layers which form columnar pores. There has been no definitive
  answer in literature regarding the number of monomers in each layer. 
%MRS3: the intro to the list talks about only ``numbers of monomers in each layer'', whereas the list below talks about several questions relating to layers. Can you rephrase so it's more general? Maybe just take out that sentence above.
We want to
  know 
\begin{enumerate}
\item If layers do exist, how many monomers constitute a single layer? \label{point:monomernum}
  \begin{itemize}
    \item A simple molecular simulation study of a similar molecule suggested
    that there are 4 monomers in each layer. Their estimation is based on a
    simulated system containing only 16 total monomers which likely does not sufficiently
    model the chemical environment present in the real system.~\cite{zhu_methacrylated_2006}. 
    \item A separate calculation based on the volume of the liquid crystal monomers proposes
    that there are seven monomers in each layer~\cite{resel_structural_2000}. 
    \item A molecular model orders of magnitude larger than any other reported atomistic 
    liquid crystal membrane simulations has the best chance of directly answering this question.
    \item We can directly change the layer composition and note its effect on membrane structure.
  \end{itemize}

@@

  \item Is it necessary to include any water in order to appropriately model the 
  Col\textsubscript{h} phase? \label{point:water}
  \begin{itemize}
	\item While the Col\textsubscript{h} phase is described as dry, it is 
          %MRS3: ``possible'' -> likely, and give some explanations why we think likely?
	possible that small amounts of ambient water may be leached into the system.
	\item The hydrogen bonding network formed by the water may play a role 
        in structuring the pore.
	\item We can use simulated X-ray diffraction patterns to see if there is
        any meaningful structural difference between a "dry" and "wet" system
  \end{itemize}
  \end{enumerate}

@@

  We developed equilibration schemes to create dry and wet configurations. Both schemes 
  start with an initial configuration generated according to the previous guidelines.
  \begin{itemize}
      \item Dry equilibration scheme:
      \begin{itemize}
        %MRS2: what initial force constant, and does it matter (that can be in supplemental) 
	%BJC: I've never explored whether there is any dependence on initial force constant. It probably doesn't matter unless it starts really low
        %MRS3: then just state what you did, so it could be repeated. 
          \item Restraints fix monomer head groups in the sandwiched or parallel-displaced
          configurations while allowing monomer tails to settle.
          \item Doing so also mitigates system dependence on initial monomer configuration.
          \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
          \item Every 50 ps, we reduce the force constants by the square root of its
          previous value, starting from 1e6 KJ mol$^-1$ nm$^-2$.
          \item Once the force constant is below 10 KJ mol$^-1$ nm$^-2$, the restraints are
          released linearly until there is no more restraining potential.
          \item The resulting unrestrained structure is allowed to equilibrate for 5 ns 
	  in the NPT ensemble with pressure controlled by the berendsen barostat.
	  \item Long, NPT equilibration simulations are run for at least 400 ns using the 
          Parrinello-Rahman barostat with a time consant of 10 ps.
      \end{itemize}
      \item Wet equilibration scheme: 
      \begin{itemize}
	  % BJC: the following three bullet points ended up not being the case since I used significantly less water
          %\item We obtained an estimate of equilibrium pore water content by solvating
          % an initial configuration with water baths above and below the membrane. 
	  %\item We ran the solvated system according to Scheme A but with a 1000 ns equilibration
          %\item Using that number we added water to the pores of an initial 
	  %configuration and equilibrated according to scheme B. 
	  \item In order to create a wet system, we solvated an initial configuration with water using gmx solvate
          \item All water molecules placed outside the pore region are removed 
	  \item Waters inside the pore region are randomly removed until the desired concentration
	  of water in the pores is reached. 
	  \item The remainder of the equilibration is the same as the dry system
      \end{itemize}
      \item In all cases, the v-rescale thermostat was used with tau-t = 0.1 % BJC: supplemental
  \end{itemize}

@@

  \begin{figure}
	\centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{tails_topview.png}  % picture of top of unit cell with only tail atoms shown
		\caption{}\label{fig:topdown_tails_only}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{tails_rzplot.png}
		\caption{}\label{fig:tails_rzplot}
	\end{subfigure}
	\caption{(a) All atoms except carbon atoms making up the tails are removed from the 
	trajectory. (b) The simulated diffraction pattern of the tail-only trajectory still 
        shows R-spots}\label{fig:tails}
        %MRS2: might need to say how the dynamic range is adjusted here with only tails, since otherwise people will be confused with the R-spots being more intense. 
	%BJC: what do you mean by dynamic range?
  \end{figure}

  \begin{figure}
  \centering
  	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
        		\centering
        		\includegraphics[width=\linewidth]{offset_tail_packing.png}
        		\caption{}~\label{fig:offset_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
		\centering
	        	\includegraphics[width=\linewidth]{offset_angle_v_I.png}
		        \caption{}~\label{fig:offset_integration}
		\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
	        \centering
		        \includegraphics[width=\linewidth]{angles_traj_layered.png}
		        \caption{}~\label{fig:rz_layered}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
        	\centering
		        \includegraphics[width=\linewidth]{layered_angle_v_I.png}
		        \caption{}~\label{fig:layered_integration}
		\end{subfigure}
	\end{subfigure}
%MRS: 
  \caption{The distribution of angles w.r.t. the xy plane between alkane chain tail centroids and nearest
  neighbor centroids for equilibrated parallel displaced (a) and sandwiched (c) configurations. The
  same peaks are visible when the 2D simulated diffraction data is radially integrated in the R-alkanes region,
  (b) and (d) respectively.}~\label{fig:tail_packing}
%MRS2: not clear what the difference is between right and left side figures.
%BJC: That's good! One is angles measured between tail centroids and the other is radial integration of the simulated xrd patterns. I'll try make that more clear in the caption.
%MRS3: sorry, I meant to say not clear what the SOURCE of the data is in right vs. left.  Make sure colors and labels are consistent in final 
%MRS: not clear why some are angles with x axis, and some are angles w neighbors?  Shouldn't it be the same thing plotted in both graphs?
  \end{figure}

>>>>

  Despite having structural data, there is still information which
  experiment cannot definitively answer. There are several key
  questions that we will investigate  
%MRS3: we can't ``answer'' for sure, since it's a model, and may omit things. Could put ``several key questions we will investigate'' or ``hope to provide evidence for or against'', or something like that.
which will be laid out and numbered
  in subsequent paragraphs.

@@

  Monomers in the Col\textsubscript{h} system are theorized to be partitioned
  into stacked layers which form columnar pores. 
%BJC: commenting out :There has been no definitive answer in literature regarding the number of monomers in each layer. 
%MRS3: the intro to the list talks about only ``numbers of monomers in each layer'', whereas the list below talks about several questions relating to layers. Can you rephrase so it's more general? Maybe just take out that sentence above.
We want to know 
\begin{enumerate}
\item If layers do exist, how many monomers constitute a single layer? \label{point:monomernum}
  \begin{itemize}
    \item A simple molecular simulation study of a similar molecule suggested
    that there are 4 monomers in each layer. Their estimation is based on a
    simulated system containing only 16 total monomers which likely does not sufficiently
    model the chemical environment present in the real system.~\cite{zhu_methacrylated_2006}. 
    \item A separate calculation based on the volume of the liquid crystal monomers proposes
    that there are seven monomers in each layer~\cite{resel_structural_2000}. 
    \item A molecular model orders of magnitude larger than any other reported atomistic 
    liquid crystal membrane simulations has the best chance of directly answering this question.
    \item We can directly change the layer composition and note its effect on membrane structure.
  \end{itemize}

@@

  \item Is it necessary to include any water in order to appropriately model the 
  Col\textsubscript{h} phase? \label{point:water}
  \begin{itemize}
	\item While the Col\textsubscript{h} phase is described as dry, it is 
          %MRS3: ``possible'' -> likely, and give some explanations why we think likely? 
	likely that small amounts of ambient water may be leached into the system.
	\item The hydrogen bonding network formed by the water may play a role 
        in structuring the pore.
	\item We can use simulated X-ray diffraction patterns to see if there is
        any meaningful structural difference between a "dry" and "wet" system
  \end{itemize}
  \end{enumerate}

@@

  We developed equilibration schemes to create dry and wet configurations. Both schemes 
  start with an initial configuration generated according to the previous guidelines.
  \begin{itemize}
      \item Dry equilibration scheme:
      \begin{itemize}
        %MRS2: what initial force constant, and does it matter (that can be in supplemental) 
	%BJC: I've never explored whether there is any dependence on initial force constant. It probably doesn't matter unless it starts really low
        %MRS3: then just state what you did, so it could be repeated. 
          \item Restraints with a force constant of 1e6 KJ mol$^-1$ nm$^-2$, fix monomer 
	  head groups in the sandwiched or parallel-displaced configurations while 
          allowing monomer tails to settle.
          \item Doing so also mitigates system dependence on initial monomer configuration.
          \item The restrained portion of the equilibration scheme is run in the NVT ensemble.
          \item Every 50 ps, we reduce the force constants by the square root of its
          previous value %, starting from 1e6 KJ mol$^-1$ nm$^-2$. <-- stated in first bullet
          \item Once the force constant is below 10 KJ mol$^-1$ nm$^-2$, the restraints are
          released linearly until there is no more restraining potential.
          \item The resulting unrestrained structure is allowed to equilibrate for 5 ns 
	  in the NPT ensemble with pressure controlled by the berendsen barostat.
	  \item Long, NPT equilibration simulations are run for at least 400 ns using the 
          Parrinello-Rahman barostat with a time consant of 10 ps.
      \end{itemize}
      \item Wet equilibration scheme: 
      \begin{itemize}
	  % BJC: the following three bullet points ended up not being the case since I used significantly less water
          %\item We obtained an estimate of equilibrium pore water content by solvating
          % an initial configuration with water baths above and below the membrane. 
	  %\item We ran the solvated system according to Scheme A but with a 1000 ns equilibration
          %\item Using that number we added water to the pores of an initial 
	  %configuration and equilibrated according to scheme B. 
	  \item In order to create a wet system, we solvated an initial configuration with water using gmx solvate
          \item All water molecules placed outside the pore region are removed 
	  \item Waters inside the pore region are randomly removed until the desired concentration
	  of water in the pores is reached. 
	  \item The remainder of the equilibration is the same as the dry system
      \end{itemize}
      \item In all cases, the v-rescale thermostat was used with tau-t = 0.1 % BJC: supplemental
  \end{itemize}

@@

  \begin{figure}
	\centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{tails_topview.png}  % picture of top of unit cell with only tail atoms shown
		\caption{}\label{fig:topdown_tails_only}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{tails_rzplot.png}
		\caption{}\label{fig:tails_rzplot}
	\end{subfigure}
	\caption{(a) All atoms except carbon atoms making up the tails are removed from the 
	trajectory. (b) The simulated diffraction pattern of the tail-only trajectory still 
        shows R-spots}\label{fig:tails}
        %MRS2: might need to say how the dynamic range is adjusted here with only tails, since otherwise people will be confused with the R-spots being more intense. 
	%BJC: what do you mean by dynamic range?
	%BJC: Explain how colorbar is adjusted
  \end{figure}

  \begin{figure}
  \centering
  	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
        		\centering
        		\includegraphics[width=\linewidth]{offset_tail_packing.png}
        		\caption{}~\label{fig:offset_tails}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
		\centering
	        	\includegraphics[width=\linewidth]{offset_angle_v_I.png}
		        \caption{}~\label{fig:offset_integration}
		\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
	\centering
		\begin{subfigure}{0.45\textwidth}
	        \centering
		        \includegraphics[width=\linewidth]{angles_traj_layered.png}
		        \caption{}~\label{fig:rz_layered}
		\end{subfigure}
		\begin{subfigure}{0.45\textwidth}
        	\centering
		        \includegraphics[width=\linewidth]{layered_angle_v_I.png}
		        \caption{}~\label{fig:layered_integration}
		\end{subfigure}
	\end{subfigure}
%MRS: 
  \caption{The distribution of angles w.r.t. the xy plane between alkane chain tail centroids and nearest
  neighbor centroids for equilibrated parallel displaced (a) and sandwiched (c) configurations. The
  same peaks are visible when the 2D simulated diffraction data is radially integrated in the R-alkanes region,
  (b) and (d) respectively.}~\label{fig:tail_packing}
%MRS2: not clear what the difference is between right and left side figures.
%BJC: That's good! One is angles measured between tail centroids and the other is radial integration of the simulated xrd patterns. I'll try make that more clear in the caption.
%MRS3: sorry, I meant to say not clear what the SOURCE of the data is in right vs. left.  Make sure colors and labels are consistent in final 
%MRS: not clear why some are angles with x axis, and some are angles w neighbors?  Shouldn't it be the same thing plotted in both graphs?
%BJC: I just need to get the labels right, they are the same thing technically
  \end{figure}
